A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Beam Dynamics

### 2018-07-11 paper/theor-experiment.tex
edits to timing and hyper parameter sections

In order to compute the peaks, we employed an additional Python script.
This script was called in the \textsc{OPAL}~input file, after the simulation
  finished using the \texttt{SYSTEM} functionality.
Once the peaks (in a given range) were located, the two objectives
  (\ref{eq:swissfel:p1}) and (\ref{eq:swissfel:p2}) were computed and their
  values written into corresponding files.
The custom \texttt{fromFile} functor allows us to access the values stored in
  the peak finder Python script result files
%
%\vspace{0.2cm}
\begin{flushleft}
\begin{Verbatim}[fontsize=\scriptsize]
rmsx:  OBJECTIVE, EXPR="statVariableAt("rms_x-err.dat", "var")";
emitx: OBJECTIVE, EXPR="statVariableAt("emit_x-err.dat", "var")";
match: OBJECTIVE, EXPR="fabs(statVariableAt("emit_x-peak.dat", "var") -
statVariableAt("rms_x-peak.dat", "var"))";
\end{Verbatim}
\end{flushleft}
\vspace{0.2cm}

\noindent
The design variables and the assembly of the multi-objective optimization problem
  can be included in the \textsc{OPAL}~input file as shown below:
%\vspace{0.2cm}
\begin{flushleft}
\begin{Verbatim}[fontsize=\scriptsize]
d1: DVAR, VARIABLE="SIGX", LOWERBOUND="0.00025", UPPERBOUND="0.00029";
d2: DVAR, VARIABLE="FIND1_MSOL10_i", LOWERBOUND="110", UPPERBOUND="120";
d3: DVAR, VARIABLE="D_LAG_RGUN", LOWERBOUND="-0.1", UPPERBOUND="0.1";
d4: DVAR, VARIABLE="D_SOLPOS", LOWERBOUND="-0.05", UPPERBOUND="0.05";

@@

%Currently we run the simulations on a low resolution space charge grid ($8
  %\times 8 \times 8$) as a proof of concept.
%In a next step a simple Python script would be applied to iteratively rerun
  %the optimizer with increasing space-charge resolution when the limits of the
  %design variables are narrow enough (iteratively diminishing search space). 
%\vspace{-7em}
\subsection{AWA Photoinjector Optimization} \label{awaproblem}
Next we apply the optimization framework to a high charge beam line
 at the Argonne Wakefield Accelerator (AWA) facility. 
The goal of this optimization is to produce beams of electrons that meet 
design specifications; this includes number of particles (charge), energy, and particle distribution (characterized by beam sizes and energy spread).
As shown in Fig.~\ref{awa-linac}, the installed portion of the the 
beam line consists of an rf photocathode gun, 
two solenoids, and six linear accelerating cavities
followed by four quadrupoles and a stripline kicker. 
The charge of interest, 40 nC, is needed for Two Beam 
Acceleration (TBA) \cite{gai_power_jing_2012,JING201872} 
experiments performed at AWA, which motivates this work. 
Prior experimental results were limited by beam size as the beam passed through small aperture 
wakefield structures located downstream. 
In an attempt to maximize charge transmission in upcoming experiments
we optimize magnet strengths in the gun and quadrupoles leading into the 
TBA section of the beam line, shown in Fig.~\ref{awa-tba}.
The simulation model includes the linac, kicker, and septum. 
The optimization location is chosen as the entrance to the fifth quadrupole, $s_3$, on the dog leg.
Minimizing beam sizes here will enable capture and further focus with quadrupoles five and six before space charge effects dominate the beam. 
This will also enable cleaner transport through downstream elements.
%\vspace{-1em}

@@

\def \septheight {65}
\def \bottomarrow {57}
\def \kleft {13}
\def \sleft {41.5}
\def \sheight {47}
\def \beam {32}
\begin{figure*}
	\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},x=1mm, y=1mm,]   
	\node (fig2) at (0,0)
	{\includegraphics[width=\textwidth]{Report/tba_dogleg}};
	\node[fill=white, inner sep=2pt] (txt2) at (3,\sheight) {\large $s_1$};
	\node[fill=white, inner sep=2pt] (txt2) at (20,\sheight) {\large $s_2$};
	\node[fill=white, inner sep=2pt] (txt2) at (46,45) {\large $s_3$};
	\node[fill=white, inner sep=2pt] (txt2) at (33,\septheight) {\large Septum};
	\draw [blue, ->, line width=1] (\sleft,\septheight) -- (\sleft, \bottomarrow);	
	\node[fill=white, inner sep=2pt] (txt2) at (7,\septheight) {\large Kicker};
	\draw [blue, ->, line width=1] (\kleft,\septheight) -- (\kleft, \bottomarrow);
	\draw [black, ->, line width=1] (0,\beam) -- (100,\beam);
	\node[fill=white, inner sep=2pt] (txt2) at (50,\beam+2) {\large Beam Direction};
	\end{tikzpicture}
	\vspace{-10em}
	\caption{Continuation of the high charge beam line layout at the AWA, top view. 
		This is the proposed two beam acceleration section. 
		Only the kicker in this drawing is installed}
	\label{awa-tba}
\end{figure*}
%
%\vspace{-1em}

In addition to the physics challenges, we chose this model to demonstrate the ability of the framework to tackle large problems.
Six design variables and objectives were used, along with three constraints.
The objectives include transverse and longitudinal beam sizes, 
transverse momentum, and longitudinal energy spread. 
The design variables include the two gun solenoids and 
the first four quadrupoles. The objectives are calculated at 
$s_3=19.4$~[m], at the entrance of the fifth quadrupole in Fig.~\ref{awa-tba}.  
This problem encompasses high dimensionality 
and nonlinear effects such as space charge. There is no existing information
that accurately predicts where the Pareto front will fall. This work is also meaningful in that
it will guide future operations at the AWA.
\nrnote{Moved final run set up and results to after time step scan and hyper parameters}

\subsubsection{Time Step Scan} \label{awa:subsection:test}
Before running a full scale optimization of the problem described in \ref{awaproblem}, 
a study on time step and number of particles in the simulation model 
was done to reduce the time to simulation while 
maintaining the physics of interest. 
The grid size $16 \times 16 \times 32$ was chosen, 
and parallelized in the x and y directions.
After comparing several options (1,000, 10,000, 20,000, 50,000, 100,000) 
with a small time step, the number of particles was fixed at ten thousand.
Next several time steps were explored, see Table~\ref{timestep}.
The largest steps were too big to resolve the physics needed.
See low fidelity plot in Fig.~\ref{tstep} for $dT=$\num{5E-11} results.  
In the drifts and linac tanks, \num{1E-11} was sufficient, 
but still not acceptable near the quadrupoles.

in combination with adjusting the time step (dT) near sensitive elements 
(i.e. quadurpoles, kicker, septum). 
The resulting simulations are low fidelity, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig.~\ref{tstep}. For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but we see discrepancies in the transverse 
(rms$_x$ and $\epsilon_x$) for low fidelity results. This discrepancy is 
what led to The
average run time of each simulation with the adjusted time steps was 1.6~minutes.
In comparison the mid fidelity simulations ran for 18~minutes.
Note, a smaller time step, \num{1E-13}, is always used in the gun where the 
beam is low energy and changing rapidly.
%A summary of the parameters used is listed in Table \ref{fidelity}.

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{Report/timestep_comparison}
	\caption{Comparison of different fidelity models (dT stands for time step).}
	\label{tstep}
\end{figure}

@@

@@

\subsubsection{Hyper parameter Scan}
While the optimization problem and goals were well defined (Section \ref{awaproblem}), 
it was not clear what the best input parameters for the NSGA-II should be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line and design variables given in Section \ref{awaproblem},
\nrnote{Changed variables for final run, the variables in 5.3 no longer match these small runs. Need to rework this section} 
four optimization experiments were done with varying NSGA-II parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
			The max number of individuals per generation was~80.}
		\label{extable}
		\begin{tabular*}{0.48\textwidth}{l @{\extracolsep{\fill}} C C D }
			\hline\noalign{\smallskip}
			& Gene Mutation Probability & Recombination Probibility & Number of completed generations \\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

The maximum number of individuals per generation was fixed at 80, 
Each experiment was allowed to run for twenty four hours, with 
a maximum generation limit of 100. 
We reduced the seven objectives to the four needed to construct 
the Pareto fronts that would be the final goal of this study.
The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)\text{, } \text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
\lsnote{You could define $\varepsilon_{x1}$ to be $\varepsilon_{x} (s=s1)$ and similarly in the table and then use going forward to simplify notation}
The optimization parameters specific to OPAL input for ex-3 are given as an example.
Note the design variables defined in Section \ref{awaproblem}
precede the optimization parameters in the OPAL input file.
%

\vspace{0.2cm}
\begin{Verbatim}[fontsize=\scriptsize]
OPTIMIZE, INPUT="tmpl/optLinac_40nC.tmpl",
OUTPUT="optLinac_40nC",
OUTDIR="results",
OBJECTIVES = {rmss1, rmsx1, emitx1, de1, rmss2, rmsx2, emitx2},
DVARS = {dv0, dv1, dv2, dv3, dv4, dv5, dv6, dv7},
CONSTRAINTS = {c1, c2},
INITIALPOPULATION=656,
MAXGENERATIONS=200,
NUM_MASTERS=1,
NUM_COWORKERS=8,
SIMTMPDIR="tmp",
TEMPLATEDIR="tmpl",
FIELDMAPDIR="fieldmaps",
NUM_IND_GEN=328,
GENE_MUTATION_PROBABILITY=0.8,
MUTATION_PROBABILITY=0.8,
RECOMBINATION_PROBABILITY=0.2;
QUIT;
\end{Verbatim}
\vspace{0.2cm}

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty four hours,
Pareto fronts at $s_1$ and $s_2$, and the design variables at the Pareto front. 
From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated about 30 to 40 fewer generations than the other experiments.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto}, we can see this is not the case.
The Pareto fronts at $s_1$, are nearly identical. We could expect
they would continue to converge given more time. 
Similar arguments can be made for ex-2, which evaluated about 15 less generations
with no clear benefit in either Pareto front.
When looking at the Pareto front at $s_2$, only ex-4 has a slight 
edge over the others.
With ex-2 and ex-3 eliminated due to evaluation time, 
the hyper parameters in ex-4 were chosen as the default values for subsequent runs.
\iffalse  
\begin{figure}
	\centering
	\includegraphics[width=0.75\textwidth]{Report/ex-pareto1}%
	\llap{\raisebox{1.8cm}{%  move next graphics to top right corner
			\includegraphics[height=5.5cm]{Report/ex-pareto2}%
	}}
	\caption{My overlay figure.}
\end{figure}  
\fi

@@

\subsubsection{TBA Optimization Problem}
\begin{align}
\text{min}  \quad & \text{rms}_{x}\left(s = s_3\right), \quad \text{rms}_{y}\left(s = s_3\right) \label{eq:awa:p1}\\
& \text{rms}_{px}\left(s = s_3\right), \quad \text{rms}_{py}\left(s = s_3\right), \label{eq:awa:p2}\\
& \text{rms}_{s}\left(s = s_3\right), \quad dE \label{eq:awa:p4} \\
\text{subject to} \quad & rms_x(s=s_1) < 0.1 \, [m]\label{eq:awa:c1}\\
\quad & rms_y(s=s_1) < 0.1\, [m]\label{eq:awa:c2}\\
\quad & |rms_y(s=s_1) - rms_x(s=s_1) | < 0.005 \, [m]\label{eq:awa:c3}\\
\quad & q = 40 \left[\text{nC}\right] \label{eq:awa:firstconstr}\\
\quad & \text{Volt}_{\text{Gun}} = 64\left[\text{MV/m}\right] \label{eq:awa:lastconstr}\\
\quad & \text{Volt}_{\text{Linac}} = 24-25\left[\text{MV/m}\right] \\
\quad & R_x = R_y = 9 \left[\text{mm}\right] \label{eq:awa:firstdvar}\\
\quad & \phi_{\text{gun}} =-20^\circ \label{eq:awa:gphidvar}\\
\quad & \phi_{\text{linac}} =-20^\circ \label{eq:awa:lastdvar}
\end{align}

>>>>

In order to compute the peaks, we employed an additional Python script.
This script was called in the \textsc{OPAL}~input file, after the simulation
  finished using the \texttt{SYSTEM} functionality.
Once the peaks (in a given range) were located, the two objectives
  (\ref{eq:swissfel:p1}) and (\ref{eq:swissfel:p2}) were computed and their
  values written into corresponding files.
The custom \texttt{fromFile} functor allows us to access the values stored in
  the peak finder Python script result files
%
%\vspace{0.2cm}

\begin{flushleft}
\begin{Verbatim}[fontsize=\scriptsize]
rmsx:  OBJECTIVE, EXPR="statVariableAt("rms_x-err.dat", "var")";
emitx: OBJECTIVE, EXPR="statVariableAt("emit_x-err.dat", "var")";
match: OBJECTIVE, EXPR="fabs(statVariableAt("emit_x-peak.dat", "var") -
statVariableAt("rms_x-peak.dat", "var"))";
\end{Verbatim}
\end{flushleft}
\vspace{0.2cm}

\noindent
The design variables and the assembly of the multi-objective optimization problem
  can be included in the \textsc{OPAL}~input file as shown below:
%\vspace{0.2cm}

\begin{flushleft}
\begin{Verbatim}[fontsize=\scriptsize]
d1: DVAR, VARIABLE="SIGX", LOWERBOUND="0.00025", UPPERBOUND="0.00029";
d2: DVAR, VARIABLE="FIND1_MSOL10_i", LOWERBOUND="110", UPPERBOUND="120";
d3: DVAR, VARIABLE="D_LAG_RGUN", LOWERBOUND="-0.1", UPPERBOUND="0.1";
d4: DVAR, VARIABLE="D_SOLPOS", LOWERBOUND="-0.05", UPPERBOUND="0.05";

@@

%Currently we run the simulations on a low resolution space charge grid ($8
  %\times 8 \times 8$) as a proof of concept.
%In a next step a simple Python script would be applied to iteratively rerun
  %the optimizer with increasing space-charge resolution when the limits of the
  %design variables are narrow enough (iteratively diminishing search space). 
%\vspace{-7em}

\subsection{AWA Photoinjector Optimization} \label{awaproblem}
Next we apply the optimization framework to a high charge beam line
 at the Argonne Wakefield Accelerator (AWA) facility. 
The goal of this optimization is to produce beams of electrons that meet 
design specifications; this includes number of particles (charge), energy, and particle distribution (characterized by beam sizes and energy spread).
As shown in Fig.~\ref{awa-linac}, the installed portion of the the 
beam line consists of an rf photocathode gun, 
two solenoids, and six linear accelerating cavities
followed by four quadrupoles and a stripline kicker. 
The charge of interest, 40 nC, is needed for Two Beam 
Acceleration (TBA) \cite{gai_power_jing_2012,JING201872} 
experiments performed at AWA, which motivates this work.

Prior experimental results were limited by beam size when the beam passed through small aperture 
wakefield structures located downstream. 
In an attempt to maximize charge transmission in upcoming experiments
we optimize magnet strengths in the gun and quadrupoles leading into the 
TBA section of the beam line, shown in Fig.~\ref{awa-tba}.
The simulation model includes the linac, kicker, and septum. 
The optimization location is chosen as the longitudinal 
entrance to the first quadrupole on the dog leg ($s_3$), see Fig.~\ref{awa-tba}.
Minimizing beam sizes here will enable capture and further focusing before space charge effects dominate the beam. 
This will also enable cleaner transport through downstream elements.
%\vspace{-1em}

@@

\def \septheight {65}
\def \bottomarrow {57}
\def \kleft {13}
\def \sleft {41.5}
\def \sheight {47}
\def \beam {35}
\begin{figure*}
	\begin{tikzpicture}[every node/.style={anchor=south west,inner sep=0pt},x=1mm, y=1mm,]   
	\node (fig2) at (0,0)
	{\includegraphics[width=\textwidth]{Report/tba_dogleg}};
	\node[fill=white, inner sep=2pt] (txt2) at (3,\sheight) {\large $s_1$};
	\node[fill=white, inner sep=2pt] (txt2) at (20,\sheight) {\large $s_2$};
	\node[fill=white, inner sep=2pt] (txt2) at (46,45) {\large $s_3$};
	\node[fill=white, inner sep=2pt] (txt2) at (33,\septheight) {\large Septum};
	\draw [blue, ->, line width=1] (\sleft,\septheight) -- (\sleft, \bottomarrow);	
	\node[fill=white, inner sep=2pt] (txt2) at (7,\septheight) {\large Kicker};
	\draw [blue, ->, line width=1] (\kleft,\septheight) -- (\kleft, \bottomarrow);
	\draw [black, ->, line width=1] (0,\beam) -- (100,\beam);
	\node[fill=white, inner sep=2pt] (txt2) at (50,\beam+2) {\large Beam Direction};
	\end{tikzpicture}
	\vspace{-12em}
	\caption{Continuation of the high charge beam line layout at the AWA, top view. 
		This is the proposed two beam acceleration section. 
		Only the kicker in this drawing is installed}
	\label{awa-tba}
\end{figure*}
%
%\vspace{-1em}

In addition to the physics challenges, we chose this model to demonstrate the ability of the framework to tackle large problems.
Six design variables and objectives were used, along with three constraints.
The objectives include transverse and longitudinal beam sizes, 
transverse momentum, and longitudinal energy spread. 
The design variables include the two gun solenoids and 
the first four quadrupoles. 
This problem encompasses high dimensionality 
and nonlinear effects such as space charge. There is no existing information
that accurately predicts where the Pareto fronts will fall. This work is also meaningful in that
it will guide future operations at the AWA.
\nrnote{Moved final run set up and results to after time step scan and hyper parameters}

\subsubsection{Time Step Scan} \label{awa:subsection:test}
Before running a full scale optimization of the problem described in Subsection \ref{awaproblem}, 
a study on time step and number of particles in the simulation model 
was done to reduce the time to simulation while 
maintaining the physics of interest. 
The grid size $16 \times 16 \times 32$ was chosen, 
and parallelized in the x and y directions.
After comparing several options (1,000, 10,000, 20,000, 50,000, 100,000) 
with a small time step, the number of particles was fixed at ten thousand.
Next several time steps were explored, see Table~\ref{timestep}.
The largest steps were too big to resolve the physics needed.
See low fidelity plot in Fig.~\ref{tstep} for $dT=$\num{5E-11} results.

@@

In the drifts and linac tanks, $dT=\num{1E-11}$ was sufficient, 
but not acceptable near the quadrupoles. 
For all models, the longitudinal parameters (rms$_s$ and energy) 
are calculated correctly, but we see discrepancies in the transverse 
(rms$_x$ and $\epsilon_x$) for low fidelity results. This discrepancy is 
what led to the decision to adjust the time steps w.r.t. beam line elements. 
In the linac and drift sections $dT=\num{1E-11}$ was used. 
A time step of $dT=\num{1E-12}$ was used near sensitive elements such as the quadrupoles, kicker, and septum.

The resulting simulations are low fidelity in most places, but closely approximate 
the mid fidelity simulations for metrics of interest, as shown in 
Fig.~\ref{tstep}.  Where we consider mid fidelity as a simulation of the 
beam line using $dT=\num{1E-12}$ everywhere. The
average run time of each simulation with the adjusted time steps was 1.6~minutes.
In comparison, the mid fidelity simulation ran for 18~minutes.
Note, a smaller time step, \num{1E-13}, is always used in the gun where the 
beam is low energy and changing rapidly.
%A summary of the parameters used is listed in Table \ref{fidelity}.

\begin{figure}%[h]
	\centering
	\includegraphics[width=0.8\linewidth]{Report/timestep_comparison}
	\caption{Comparison of different fidelity models (dT stands for time step).}
	\label{tstep}
\end{figure}

@@

\subsubsection{Hyper parameter Scan}
While the optimization problem and goals were well defined (Subsection \ref{awaproblem}), 
it was not clear what the best hyper parameters for the NSGA-II would be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line in Fig.~\ref{awa-linac},
four small optimization experiments were done with varying NSGA-II parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
			The max number of individuals per generation was~80.}
		\label{extable}
		\begin{tabular*}{0.48\textwidth}{l @{\extracolsep{\fill}} C C D }
			\hline\noalign{\smallskip}
			& Gene Mutation Probability & Recombination Probibility & Number of completed generations \\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

The maximum number of individuals per generation was fixed at 80, 
Each experiment was allowed to run for twenty four hours, with 
a maximum generation limit of 100. 
We reduced the six objectives to four, 
and shortened the simulation time by moving the objectives further 
upstream to $s_1$ and $s_2$.
The objectives include: $\varepsilon_{x}\left(s = s_1\right)\text{, } \varepsilon_{x}\left(s = s_2\right)\text{, } \text{rms}_{s}\left(s = s_1\right)\text{, and }  \text{rms}_{s}\left(s = s_2\right)$. 
\lsnote{You could define $\varepsilon_{x1}$ to be $\varepsilon_{x} (s=s1)$ and similarly in the table and then use going forward to simplify notation}
The OPAL input file for ex-3 is given as an example to show how optimization and design variables are defined:
%

%\vspace{0.2cm}
\begin{Verbatim}[fontsize=\scriptsize]
OPTION, ECHO=FALSE;
OPTION, INFO=TRUE;

TITLE, STRING="ANL Optimisation";

dv0: DVAR, VARIABLE="IBF",    LOWERBOUND=200.0, UPPERBOUND=500.0;
dv1: DVAR, VARIABLE="IM",     LOWERBOUND=170.0, UPPERBOUND=260.0;
dv2: DVAR, VARIABLE="GPHASE", LOWERBOUND=-30.0, UPPERBOUND=0.0;
dv3: DVAR, VARIABLE="FWHM",   LOWERBOUND=1.5,   UPPERBOUND=10.0;

// Quad values
dv4: DVAR, VARIABLE="KQ1", LOWERBOUND=-8.0, UPPERBOUND=8.0;
dv5: DVAR, VARIABLE="KQ2", LOWERBOUND=-8.0, UPPERBOUND=8.0;
dv6: DVAR, VARIABLE="KQ3", LOWERBOUND=-8.0, UPPERBOUND=8.0;
dv7: DVAR, VARIABLE="KQ4", LOWERBOUND=-8.0, UPPERBOUND=8.0;

rmss1:  OBJECTIVE,EXPR="fabs(statVariableAt('rms_s',16.5))";
emitx1: OBJECTIVE,EXPR="fabs(statVariableAt('emit_x',16.5))";

rmss2:  OBJECTIVE,EXPR="fabs(statVariableAt('rms_s',18.5))";
emitx2: OBJECTIVE,EXPR="fabs(statVariableAt('emit_x',18.5))";

c1: CONSTRAINT, EXPR="fabs(statVariableAt('rms_x',16.5))<1.0e-1";
c2: CONSTRAINT, EXPR="fabs(statVariableAt('rms_y',16.5))<1.0e-1";

OPTIMIZE, INPUT="tmpl/optLinac_40nC.tmpl",
OUTPUT="optLinac_40nC",
OUTDIR="results",
OBJECTIVES = {rmss1, emitx1, rmss2, emitx2},
DVARS = {dv0, dv1, dv2, dv3, dv4, dv5, dv6, dv7},
CONSTRAINTS = {c1, c2},
INITIALPOPULATION=80,
MAXGENERATIONS=100,
NUM_MASTERS=1,
NUM_COWORKERS=8,
SIMTMPDIR="tmp",
TEMPLATEDIR="tmpl",
\end{Verbatim}
%\vspace{0.2cm}

After collection of the data for all four experiments, several metrics
were compared, including number of generations completed in twenty four hours and
Pareto fronts at $s_1$ and $s_2$.
From Table \ref{extable}, we clearly see ex-3 is significantly 
slower, as it evaluated only 53 generations compared to the max 
experiment, ex-1 at 96 generations.
Perhaps this trade off would be acceptable if the Pareto front was significantly 
improved, but from Fig. \ref{expareto}, we can see this is not the case.
Similar arguments can be made for ex-2, which evaluated about 15 less generations
with no clear benefit in either Pareto front.
The Pareto fronts at $s_1$, are nearly identical. We could expect
this trend would continue given more time. 
When looking at the Pareto front at $s_2$, only ex-4 has a slight 
edge over the others.
With ex-2 and ex-3 eliminated due to evaluation time, 
and a slightly better Pareto front at $s_2$ for ex-4, the hyper parameters in ex-4 were chosen as the default values for subsequent runs.
\iffalse  
\begin{figure}
	\centering
	\includegraphics[width=0.75\textwidth]{Report/ex-pareto1}%
	\llap{\raisebox{1.8cm}{%  move next graphics to top right corner
			\includegraphics[height=5.5cm]{Report/ex-pareto2}%
	}}
	\caption{My overlay figure.}
\end{figure}  
\fi

@@

\subsubsection{TBA Optimization Problem}
With computational and hyper parameters set, 
the full optimization problem of interest is explored.
The objectives (beam sizes and energy spread) are calculated at 
$s_3=19.4$~[m], see Fig.~\ref{awa-tba}.

\begin{align}
\text{min}  \quad & \text{rms}_{x}\left(s = s_3\right), \quad \text{rms}_{y}\left(s = s_3\right) \label{eq:awa:p1}\\
& \text{rms}_{px}\left(s = s_3\right), \quad \text{rms}_{py}\left(s = s_3\right), \label{eq:awa:p2}\\
& \text{rms}_{s}\left(s = s_3\right), \quad dE \label{eq:awa:p4} \\
\text{subject to} \quad & rms_x(s=s_1) < 0.1 \, [m]\label{eq:awa:c1}\\
\quad & rms_y(s=s_1) < 0.1\, [m]\label{eq:awa:c2}\\
\quad & |rms_y(s=s_1) - rms_x(s=s_1) | < 0.005 \, [m]\label{eq:awa:c3}\\
\quad & q = 40 \left[\text{nC}\right] \label{eq:awa:firstconstr}\\
\quad & \text{Volt}_{\text{Gun}} = 64\left[\text{MV/m}\right] \label{eq:awa:lastconstr}\\
\quad & \text{Volt}_{\text{Linac}} = 24-25\left[\text{MV/m}\right] \\
\quad & R_x = R_y = 9 \left[\text{mm}\right] \label{eq:awa:firstdvar}\\
\quad & \phi_{\text{gun}} =-20^\circ \label{eq:awa:gphidvar}\\
\quad & \phi_{\text{linac}} =-20^\circ \label{eq:awa:lastdvar}
\end{align}
